---
title: "Pace Count Analysis"
author: "Peraton, Columbus (version 11/17/2020)"
date: "Purpose: (1) review key insights and supporting data visualizations; (2) discuss a project way-ahead."
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(Rmisc)
library(foreign)
library (memisc)
library(tidyverse)
library(readxl)
library(eRm)
library(ggcorrplot)
library(corrplot)
library(chron)
library(hms)
library(mclust)
library(imputeMissings)
library(randomForest)
library(plotly)
library(ggdendro)
library(cluster)
library(wordcloud)
library(wordcloud2)
library(stringr)
library(tm)
library(gganimate)
library(pracma)
library(lubridate)
library(tidyverse)
library(readxl)
library(corrplot)
library(janitor)
library(lubridate)
library(chron)
library(rpart)
library(rpart.plot)
library(naniar)
#library(ggstatplot)

#set dataframe and filter out ALC
df_pre <- read_xlsx("PaceCountData_1Sept.xlsx") %>%
  rbind(read_xlsx("PaceCountData_13OCT.xlsx") %>%
          mutate(duration=duration_min+(duration_sec/60)) %>% 
          select(-duration_min, -duration_sec)) %>%
  rbind(read_xlsx("PaceCountData_19Oct.xlsx") %>% 
          mutate(duration=duration_min+(duration_sec/60)) %>% 
          select(-duration_min, -duration_sec)) %>% 
  rbind(read_xlsx("PaceCountData_09NOV.xlsx") %>% 
          mutate(duration=duration_min+(duration_sec/60)) %>% 
          select(-duration_min, -duration_sec)) %>%
  mutate(duration=round(duration, 1)) %>%
  drop_na(pacecount)  %>%
  filter(course!="ALC")  %>% #remove ALC
  mutate(NVG = if_else(course =="RSLC_Cadre" & day_night=="night"|course == "RSLC_Students" & day_night=="night", "Yes", "No")) %>% 
  mutate(duration = if_else(course=="RSLC_Cadre" & day_night=="night" & loadcarrying=="load", 0, duration)) %>% 
  mutate(duration = ifelse(duration==0,NA,duration)) 

df <- df_pre %>% 
  right_join(
    df_pre %>%  
  filter(loadcarrying=="noload", day_night =="night", length=="long") %>% 
  mutate_at(vars(duration), scale)  %>% 
  rbind(
    df_pre %>%  
  filter(loadcarrying=="noload", day_night =="day", length=="long") %>% 
  mutate_at(vars(duration), scale)  
  ) %>% 
  select(ID, day_night, duration) %>% 
  pivot_wider(names_from = "day_night", values_from = "duration")  %>% 
  group_by(ID) %>% 
  mutate(duration = if_else(is.na(night), day, night )) %>% 
  mutate(ExperienceLevel = if_else(duration <= -.3, "advanced", if_else(duration<.8, "intermediate", "beginner"))) %>% 
  mutate(ExperienceLevel =  replace_na(ExperienceLevel, "intermediate")) %>% 
  select(ID, ExperienceLevel) 
  ) %>% 
  mutate(loadcarrying = factor(loadcarrying, levels = c("noload", "load"))) %>% 
  mutate(gender = factor(gender, levels=c("M", "F"))) %>% 
  mutate(day_night = factor(day_night, levels=c("day", "night"))) %>% 
  mutate(NVG = factor(NVG, levels=c("No", "Yes"))) %>% 
  mutate(ExperienceLevel= factor(ExperienceLevel, levels=c("beginner", "intermediate", "advanced"))) 
  

#summary statistics (course, gender, experience)
df  %>% group_by(ID, course, gender) %>% select(ID, course, gender) %>% unique() %>% group_by(course, gender) %>% summarise(n=n()) %>%  pivot_wider(names_from = "gender", values_from = "n")

df  %>% select(ID, course, ExperienceLevel) %>% unique() %>% group_by(course,  ExperienceLevel) %>% summarise(n=n()) %>%  pivot_wider(names_from = "ExperienceLevel", values_from = "n")

df  %>% select(ID, gender, ExperienceLevel) %>% unique() %>%  
  group_by(gender, ExperienceLevel) %>% 
  summarise(n = n()) %>% 
  pivot_wider(names_from = "ExperienceLevel", values_from = "n")

df  %>% filter(ruckweight>0) %>% select(ID, ruckweight, course) %>% unique() %>%  
  group_by(course) %>% 
  summarise(ruckweight = mean(ruckweight)) %>% 
  pivot_wider(names_from = "course", values_from = "ruckweight")

############################### Write total dataset
#write.csv(df, "PaceCountData_21Oct20.csv")

############################### Filter out the outliers
#Men outliers
Q_M_day <- df %>% filter(gender=="M", day_night=="day") %>%
     select(pacecount) %>%  quantile(probs=c(.25,.75), na.rm=TRUE)
iqr_M_day <- df %>% filter(gender=="M", day_night=="day") %>%
     select(pacecount) %>% na.omit()
iqr_M_day <- IQR(iqr_M_day$pacecount)
up_M_day <- round(Q_M_day[2]+1.5*iqr_M_day,0)
low_M_day <- round(Q_M_day[1]-1.5*iqr_M_day,0)

Q_M_night <- df %>% filter(gender=="M", day_night=="night") %>% select(pacecount) %>%  quantile(probs=c(.25,.75), na.rm=TRUE)
iqr_M_night <- df %>% filter(gender=="M", day_night=="night") %>% select(pacecount) %>% na.omit()
iqr_M_night <- IQR(iqr_M_night$pacecount)
up_M_night <- round(Q_M_night[2]+1.5*iqr_M_night,0)
low_M_night <- round(Q_M_night[1]-1.5*iqr_M_night,0)

#Female outliers
Q_F_day <- df %>% filter(gender=="F", day_night=="day") %>% select(pacecount) %>%  quantile(probs=c(.25,.75), na.rm=TRUE)
iqr_F_day <- df %>% filter(gender=="F", day_night=="day") %>% select(pacecount) %>% na.omit()
iqr_F_day <- IQR(iqr_F_day$pacecount)
up_F_day <- round(Q_F_day[2]+1.5*iqr_F_day,0)
low_F_day <- round(Q_F_day[1]-1.5*iqr_F_day,0)

Q_F_night <- df %>% filter(gender=="F", day_night=="night") %>% select(pacecount) %>%  quantile(probs=c(.25,.75), na.rm=TRUE)
iqr_F_night <- df %>% filter(gender=="F", day_night=="night") %>% select(pacecount) %>% na.omit()
iqr_F_night <- IQR(iqr_F_night$pacecount)
up_F_night <- round(Q_F_night[2]+1.5*iqr_F_night,0)
low_F_night <- round(Q_F_night[1]-1.5*iqr_F_night,0)

df_outliersremoved_F_day <- df %>% filter(gender=="F", day_night=="day") %>% 
        filter(pacecount >= low_F_day & pacecount <= up_F_day)
df_outliersremoved_F_night <- df %>% filter(gender=="F", day_night=="night") %>% 
        filter(pacecount >= low_F_night & pacecount <= up_F_night)
df_outliersremoved_M_day <- df %>% filter(gender=="M", day_night=="day") %>% 
        filter(pacecount >= low_M_day & pacecount <= up_M_day)
df_outliersremoved_M_night <- df %>% filter(gender=="M", day_night=="night") %>% 
        filter(pacecount >= low_M_night & pacecount <= up_M_night)

df_outliersremoved <- df_outliersremoved_F_day %>% 
     rbind(df_outliersremoved_F_night) %>% 
     rbind(df_outliersremoved_M_day) %>% 
     rbind(df_outliersremoved_M_night) %>%
  mutate( ExperienceLevel = factor(ExperienceLevel, levels=c("advanced", "intermediate", "beginner"))) 

#difference between day and night pacecounts by experience level
df_outliersremoved  %>% filter(length=="long", loadcarrying=="noload") %>% 
  select(ID, ExperienceLevel, gender, pacecount, day_night) %>% 
  pivot_wider(names_from = "day_night", values_from = "pacecount") %>% 
  mutate(delta = (night-day)) %>% 
  summarySE(groupvars = c("ExperienceLevel"), measurevar = "delta", na.rm = TRUE) 

################# Write file with outliers removed
#write.csv (df_outliersremoved, "PaceCountData_21Oct20_outliersremoved.csv")

```

## Sample Summary

<div class="blue">
Study sample size: Females n=`r count2_F$n`; Males n=`r count2_M$n`; ~1,500 observations.
</div>
```{r sample summary,  message=FALSE, warning=FALSE}
df %>% select (ID, course, gender) %>% unique() %>%  group_by(course, gender) %>% summarise(n=n()) %>%
  ggplot(aes(x=reorder(course, n, fun=max), y=n, fill=gender)) + 
  geom_col() + 
  ylab("# of study participants") +
  xlab("")+
  geom_text(aes(label=n), hjust=1) +
  scale_fill_manual(values=c("skyblue", "tomato")) +
  coord_flip()
```

## Key Findings (1 of 2)
 
- ***At Night*** condition has a marked impact on pace count (approx. +4 steps for advanced trainees and +10 steps for beginners).

- ***Height*** displays a moderate negative correlation with pace count (-0.4), suggesting an impact from stride length. 

- ***Load/Ruck weight*** has a marked impact to pace count (~ +4 steps from comparable noload conditions); ruck weights over 35 lbs having increasing impact.

- ***Duration*** appears to be a proxy measure for ***experience***; lower duration, especially at night, suggests greater experience and generally lower pace counts.

## Key Findings (2 of 2)

- ***Gender*** appears to have an impact on pace count, though multi-collinearity with height and "experience" reduces the information gain from this feature. 

- ***Soldier Weight*** strongly correlates with height and has negligible information gain for predicting pace count. 

- ***Night Vision Goggles*** may contribute to a minor reduction (~ -2 steps) in pace counts at night, when compared to trainees without NVGs.  This finding requires further exploration to determine if experience of RSLC Students is confounding.

- ***For further exploration:*** 20% (52/259) record lower pace counts under load, generally at night.  

## Correlation plot

```{r correlation, message=FALSE, warning=FALSE}
df_corrplot <- df_outliersremoved %>% filter(length == "long") %>% 
        select (pacecount, ruckweight, day_night, height, weight, gender,  ExperienceLevel) %>%
               mutate(day_night = if_else(day_night == "day", 0, 1), 
               gender = if_else(gender == "F", 0,1),
               experience = if_else(ExperienceLevel=="advanced", 2,
                                    if_else(ExperienceLevel=="intermediate", 1, 0))) %>%
  select(-ExperienceLevel)

corrplot(cor(df_corrplot), method="color", order="hclust", type="lower", addrect=0, cl.lim=c(-1,1), addCoef.col="black", rect.col="green", diag = FALSE, number.font=1, number.digits = 2, number.cex = .8, tl.cex = .8, cl.pos = "n", tl.srt = 45)
```

## Day/Night & Solider Height

```{r heightimpact,  message=FALSE, warning=FALSE}
df_outliersremoved %>%  filter(length=="long") %>% 
     ggplot(aes(x=height,y=pacecount, color=day_night)) +
     geom_density2d(alpha=.8) + geom_smooth(size = 2) +
     scale_color_manual(values = c("darkgray", "blue")) +
     theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
     ylab("pace count") + xlab("Soldier height (inches)") +  xlim(58,76) + ylim(55,95) +
  ggsave("day_night_height.png", height = 4 , width = 6.5, units = "in")
```

## Load (1 of 2)
- points above dashed line represent pace count ***increase***
```{r load1,  message=FALSE, warning=FALSE}
df_outliersremoved %>% 
  filter(length=="long", day_night=="day") %>%
  select(course , ID, day_night, loadcarrying, ExperienceLevel,pacecount)  %>% 
  pivot_wider(names_from=loadcarrying, values_from=pacecount) %>%
  mutate(delta = (load - noload)) %>% 
  rbind(
    df_outliersremoved %>% 
  filter(length=="long", day_night=="night") %>%
  select(course , ID, day_night, loadcarrying, ExperienceLevel, pacecount)  %>% 
  pivot_wider(names_from=loadcarrying, values_from=pacecount) %>%
  mutate(delta = (load - noload))
    ) %>%
#  summarySE(groupvars = c("day_night"), measurevar = "delta", na.rm=TRUE)
  drop_na(delta) %>% 
     ggplot(aes(x=delta, fill=day_night)) + 
     geom_density(alpha = .3) + 
        ylab("density") + 
        xlab("pace count delta under load") +
        scale_fill_manual(values = c("green", "blue")) +
  geom_vline(xintercept = 0, linetype="dashed")
        
```


## Load Carrying (Gender Differences)

```{r loadcarrying_gender, message=FALSE, warning=FALSE}

loadcarrying <- df_outliersremoved %>% 
  filter(length=="long", day_night=="day") %>%
  select(course , ID, day_night, height, gender, loadcarrying, ExperienceLevel,pacecount)  %>% 
  pivot_wider(names_from=loadcarrying, values_from=pacecount) %>%
  mutate(delta = (load - noload)) %>% 
  rbind(
    df_outliersremoved %>% 
  filter(length=="long", day_night=="night") %>%
  select(course , ID, day_night, height, gender,  loadcarrying, ExperienceLevel, pacecount)  %>% 
  pivot_wider(names_from=loadcarrying, values_from=pacecount) %>%
  mutate(delta = (load - noload)) 
  ) %>% 
    left_join(
      df_outliersremoved %>% select(ID, ruckweight) %>% filter(ruckweight>0) %>% unique()) %>% 
  ungroup() %>% 
  drop_na(delta) %>% 
  select(-load, -noload, -ID, -course) 

loadcarrying %>% mutate(category = if_else(ruckweight<=35, "light", "heavy")) %>% 
  summarySE(groupvars = c("category","gender", "day_night"), measurevar = "delta") %>% 
  arrange(gender)

loadcarrying %>% 
  ggplot(aes(x=ruckweight, y = delta, color=gender)) +
  geom_point() +
  geom_smooth() +
  xlim(30,55)
```

## Ruck Weight
  
```{r ruckweight1,  message=FALSE, warning=FALSE}
#impact of rucks > 35 lbs on pace count
#df2b %>% mutate(WeightCategory = ifelse(ruckweight>35, "heavy", "light"))  %>%  summarySE(groupvars = c("WeightCategory") , measurevar = "LoadDelta", na.rm = TRUE)
#correlation
#cor(df2b  %>% select(ruckweight, LoadDelta) %>% na.omit())

df2b %>% 
     ggplot(aes(x=ruckweight,y=LoadDelta, color=day_night)) +
     geom_density(aes(x=ruckweight, y=..count..), color="darkgray",  fill="skyblue", alpha=0.3) +
     geom_point(alpha=.5) + 
     geom_smooth() +  
     geom_hline(yintercept = 0, linetype="dashed", color="black") + 
     scale_color_manual(values = c("green", "blue")) +
     ylab("pace count delta") + 
     ylim(-10, 24) + 
     xlim(15,55) +
    xlab("ruck weight (lbs)") +
     theme(legend.title= element_text(color="black", size=10), legend.position = "right")+
   #  labs(caption = "Note: density function drawn for number of observations") +
    ggsave("ruckweight.png", height = 4 , width = 6.5, units = "in")
```

## Duration (1 of 3)

```{r durationanalysis1,  message=FALSE, warning=FALSE}

df_outliersremoved$day_night <- factor(df_outliersremoved$day_night, levels = c("night", "day"))
df_outliersremoved$loadcarrying <- factor(df_outliersremoved$loadcarrying, levels = c("noload", "load"))

df_outliersremoved %>%  filter (duration>0) %>% ggplot(aes(x=duration, y=pacecount)) +
     geom_point(aes(color=ExperienceLevel)) +
     facet_grid(day_night~.) +
  xlab("time duration (min)") +
  ylab("pace count") +
     geom_smooth() +
     labs(caption = "green dashed = mean DAY duration; red dashed = mean NIGHT duration")
```

## Duration (2 of 3)
Duration used to inform a dummy variable for experience.

```{r durationanalysis2,  message=FALSE, warning=FALSE}
df_outliersremoved %>% 
  mutate( ExperienceLevel = factor(ExperienceLevel, levels=c("advanced", "intermediate", "beginner"))) %>%
  mutate(day_night = factor(day_night, levels=c("day", "night"))) %>% 
     ggplot(aes(x=loadcarrying, y=pacecount, fill=day_night)) +
     geom_boxplot(alpha=.5) +
     stat_summary(aes(label=round(..y..,2)), fun.y="median", geom="text", size=3.5, color="red", vjust=-.2, position = position_dodge(width = .8)) +
     facet_grid(.~ExperienceLevel) +
     xlab("") +
     scale_fill_manual(values = c("darkgray", "skyblue", "tomato")) +
  theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
     ylim(55,95) +
  ylab ("pace count") +
  #geom_hline(yintercept = c(60, 70, 80, 90), linetype = "dashed", color = "gray") +
  ggsave("experience.png", height = 4 , width = 6.5, units = "in")
```

## Duration (3 of 3)

```{r durationanalysis3,  message=FALSE, warning=FALSE}
#Analysis of Impact of Proficiency.
df_outliersremoved %>% filter(loadcarrying=="noload", length=="long") %>% 
  select(ID, ExperienceLevel, day_night,pacecount) %>% 
  spread(key=day_night, value=pacecount) %>% 
  na.omit() %>% 
  mutate(day_night_noload_delta=night-day) %>% 
  select(ID, day_night_noload_delta, ExperienceLevel) %>% 
  left_join(
    df %>% filter(loadcarrying=="load", length=="long") %>% 
      select(ID, course, day_night,pacecount) %>% 
      spread(key=day_night, value=pacecount) %>% 
      na.omit() %>% 
      mutate(day_night_load_delta=night-day) %>%
      select(ID, day_night_load_delta) 
      ) %>% 
        ggplot(aes(x=day_night_noload_delta, y=day_night_load_delta, color=ExperienceLevel)) + 
        geom_density2d(size=.5)  +
        ylim(-5,25) +
        xlim(-5,25) +
        geom_abline(slope = 1, linetype="dashed", color="black", size=.5) +
        scale_color_manual("Experience level", values = c("green", "blue", "red")) +
        xlab("Day_Night pacecount delta (noload)") +
        ylab("Day_Night pacecount delta (load)") +
        theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
  ylab("pace count delta (under load)") +
  xlab("pace count delta (no load)") +
       # facet_wrap(course~.) +
  #labs(caption = "advanced = fastest 24% duration; beginner = slowest 23% duration; assessed from noload condition") +
  ggsave("experience_deltas.png", height = 4 , width = 6.5, units = "in")

```

## Gender

```{r gender,  message=FALSE, warning=FALSE}
df_outliersremoved %>%   
  ggplot(aes(x=height, y=pacecount, color=gender, group=gender)) + 
  geom_density2d(alpha=.8) + geom_smooth(size = 2) +
  ylab("pace count") +
  scale_color_manual(values = c("skyblue", "tomato")) +
   theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
  xlab("Soldier height (inches)") + 
  ggsave("gender.png", height = 4 , width = 6.5, units = "in")
```

## Solider Weight

```{r weightimpact,  message=FALSE, warning=FALSE}
df_outliersremoved %>%   mutate(wt_ht=weight/height) %>%  
     ggplot(aes(x=weight,y=pacecount)) +
     geom_density2d(alpha=.8) + geom_smooth(size = 2) +
     geom_hline(yintercept = 0, linetype="dashed", color="red") + 
     scale_color_manual(values = c("tomato", "skyblue")) +
     theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
     ylab("pace count") + xlab("weight to height ratio (lbs/inch)") +
        ylim(55,90)
```

## Night Vision

```{r nightvision1,  message=FALSE, warning=FALSE}
df_outliersremoved %>% filter(day_night=="night") %>% 
        ggplot(aes(x=height, y=pacecount, color=NVG)) +
        geom_point(alpha=.5) +
  ylab("pace count") +
  scale_color_manual(name = "night vision", values = c("darkgray", "blue")) +
  theme(legend.title= element_text(color="black", size=10), legend.position = "right") +
  xlim(63,76) + ylim(60,100) +
  xlab("Soldier height (inches)") +
        geom_smooth() +
  ggsave("nightvision.png", height = 4 , width = 6.5, units = "in")
```

## Precision
```{r experience2,  message=FALSE, warning=FALSE}
df_precision_plot <- df %>% select (course, ID, pacecount) %>% group_by(course, ID) %>% summarise(pacecount_sd = sd(pacecount)) 

df_precision_plot %>% 
  ggplot(aes(x=reorder(course, pacecount_sd,fun=median), y=pacecount_sd)) + 
  geom_boxplot() +
  stat_summary(aes(label=round(..y..,1)), fun.y="median", geom="text", size=3, color="red", vjust=-.2) + ylim(0,15) + xlab("") + ylab ("participant pace count std deviation") +
  labs(caption = "Note: std deviation determined from 4 trials (night/day, load/noload)")
```

## What Matters (1 of 3) |Multiple Linear Regression

***Essential Features (R2 = 0.535):***

- Day or Night Condition
- Soldier Height  
- Load/Ruck weight or No load Condition 

***Added Value (R2 = 0.597):***

- Experience (beginner/intermediate/advanced) 

***Small Information Gain:***

- Gender (R2 = 0.614)  
- Soldier Weight (R2 = 0.615) 

## What Matters (2 of 3) |Machine Learning (Regression Trees)
```{r whatmatters1,  message=FALSE, warning=FALSE}
m <- rpart (pacecount ~ ExperienceLevel + height + day_night + ruckweight , data = df_outliersremoved %>%  
     filter(length=="long"))
predict_trees <- predict(m, df_regressiontrees)
regtree_importance <- as.data.frame(m$variable.importance) %>% as.data.frame()
max_importance <- max(regtree_importance$`m$variable.importance`)
regtree_report <- rownames_to_column(regtree_importance, var="Feature") %>%
        mutate(Importance = round(`m$variable.importance`/max_importance,1)) %>% 
        select ( Feature, Importance) %>%  arrange(-Importance)
regtree_report %>% ggplot () + geom_col(aes(x=reorder(Feature, Importance, fun=max), y=Importance), fill= "skyblue") + coord_flip() + xlab("") + ylab("importance in pace count prediction")
```

## What Matters (3 of 3)|Machine Learning (Regression Trees)

```{r whatmatters2,  message=FALSE, warning=FALSE}
rpart.plot(m, digits=2, fallen.leaves = FALSE, type=5, extra=0, tweak = 1.1)
```

## Course Comparison (1 of 2)
Short (100m) vs. long (600m) course results: points expected near dashed line

```{r coursecomparison2,  message=FALSE, warning=FALSE}
df_outliersremoved %>% filter(loadcarrying=="noload", day_night=="day") %>% 
        select(course, ID, ExperienceLevel, length, pacecount) %>% 
        spread(key=length, value=pacecount)  %>% 
        na.omit() %>%
        mutate(delta=(long-short)) %>%
  ggplot(aes(x=delta)) + geom_histogram(binwidth = 2)
  
df_outliersremoved %>% filter(loadcarrying=="noload", day_night=="day") %>% 
        select(course, ID, ExperienceLevel, length, pacecount) %>% 
        spread(key=length, value=pacecount)  %>% 
        na.omit() %>%
        mutate(delta=(long-short)) %>% 
  mutate(precise = if_else(abs(delta) <= 5, "yes", "no")) %>%
  mutate(direction = if_else(delta>2, "long", if_else(delta<(-2), "short", "none"))) %>% 
  summarySE(groupvars = "precise", measurevar = "delta")

df_outliersremoved %>% filter(loadcarrying=="noload", day_night=="day") %>% 
        select(course, ID, ExperienceLevel, length, pacecount) %>% 
        spread(key=length, value=pacecount)  %>% 
        na.omit() %>%
        mutate(Delta=abs(long-short)) %>% 
        mutate(Delta_Courses=ifelse(Delta>3, "large(>3)", "small(<=3)")) %>% 
     ggplot(aes(x=short, y=long)) +
     geom_density_2d(alpha=.7) + 
     geom_abline(slope = 1, linetype="dashed", color='red') + 
     scale_color_manual(values=c("tomato", "skyblue")) +
     theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
     ylab("pace count (600m course)") + 
     xlab("pace count (100m course)") +
     xlim(55,85) + 
     ylim(55,85) + facet_grid(.~ExperienceLevel)
```

## Course Comparison (2 of 2)

```{r course comparison1, message=FALSE, warning=FALSE}
#Comparison of short and long course results (day, noload)
df_outliersremoved %>% 
     filter(loadcarrying=="noload", day_night=="day") %>% 
     ggplot(aes(x=height, y=pacecount, color=length)) + geom_point() + 
     geom_smooth() +
     scale_color_manual(values=c("tomato", "green")) +
     theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
         labs(caption = "Note: day and noload conditions")
```

## Study Comparison |Ranger/RSLC Land Nav vs. Pace Count Study

```{r studycomparison, message=FALSE, warning=FALSE}
#Comparison of study results
df_LandNav <- read_xls("LandNavResults_4Nov20.xls") %>% 
  select(WalkingPaceCount, NightPaceCount2, RuckPaceCount2) %>% 
  gather(WalkingPaceCount:RuckPaceCount2, key=Category, value=Pacecount) %>% 
  na.omit() %>% 
  mutate(Event="LandNavStudy")
df_LandNav <- rename(df_LandNav, pacecount = "Pacecount")

a <- df_outliersremoved %>% 
  filter(day_night=="day", loadcarrying=="noload", gender=="M") %>% mutate(Event="PaceCountStudy", Category="WalkingPaceCount") %>% 
  select(Category, pacecount, Event)  
b <- df_outliersremoved %>% 
  filter(day_night=="night", loadcarrying=="noload", gender=="M") %>% 
  mutate(Event="PaceCountStudy", Category="NightPaceCount2") %>% 
  select(Category, pacecount, Event)  
c <- df_outliersremoved %>% 
  filter(day_night=="day", loadcarrying=="load", gender=="M") %>%
  mutate(Event="PaceCountStudy", Category="RuckPaceCount2") %>% 
  select(Category, pacecount, Event)

df_comparison <- df_LandNav %>% rbind(a, b, c) %>%
  mutate(pacecount = as.numeric(pacecount) )  %>%
  na.omit(pacecount) %>% filter(pacecount<110)

df_comparison %>% summarySE(measurevar = "pacecount", groupvars = c("Event", "Category")) %>% 
     ggplot(aes(x=reorder(Category, pacecount, fun=mean), y=pacecount, color=Event, group=Event)) +
     geom_point(size=3)+
     geom_line() + xlab("")+
     geom_errorbar(aes(ymin=pacecount-ci, ymax=pacecount+ci), width=.1 ) +
     geom_text(aes(label=round(pacecount,1)), size=4, vjust=-.4, color="black") +
     theme(legend.title= element_text(color="black", size=10), legend.position = "top")  + ylab("Pace Count Group Mean") +
     ylim(65,75) + labs(caption = "Note: sample data includes males only; 62% of Ranger Students did not report a night pace count")
```

## Pace Count Table

Table available at:

https://github.com/mattdeloia/PaceCount_Analysis/blob/gh-pages/PaceCountTable(Final).png

Interactive Application (under development) at:
https://mattdeloia.shinyapps.io/PaceCount_Estimator/

Notes:

* Advanced Navigators make the following adjustments at each no load and load pace count: (-2) steps to all Day ; (-5) steps to all Night.
* Beginner Navigators make the following adjustments at each no load and load pace count:  +2 steps to all Day; +5 steps to all Night.

## Way-Ahead

Transition key findings for doctrine and POI/lesson plans (DOTD brief):

- Utility of a 600M cross-country pace count course
- Day versus night pace counts
- Effect of load on pace counts
- Other considerations- NVDs, thick vegetation, weather 
- Pace count estimator app?

Article submission on Dead Reckoning with results from future ***Drift Course Study.***
