---
title: "Pace Count Analysis"
author: "Northrop Grumman, Columbus"
date: "11/13/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(Rmisc)
library(foreign)
library (memisc)
library(tidyverse)
library(readxl)
library(eRm)
library(ggcorrplot)
library(corrplot)
library(chron)
library(hms)
library(mclust)
library(imputeMissings)
library(randomForest)
library(plotly)
library(ggdendro)
library(cluster)
library(wordcloud)
library(wordcloud2)
library(stringr)
library(tm)
library(gganimate)
library(pracma)
library(lubridate)
library(tidyverse)
library(readxl)
library(corrplot)
library(janitor)
library(lubridate)
library(chron)
library(rpart)
library(rpart.plot)
library(naniar)
#library(ggstatplot)

dfa <- read_xlsx("PaceCountData_1Sept.xlsx")
dfb  <- read_xlsx("PaceCountData_13OCT.xlsx") %>% mutate(duration=duration_min+(duration_sec/60)) %>% select(-duration_min, -duration_sec)
dfc  <- read_xlsx("PaceCountData_19Oct.xlsx") %>% mutate(duration=duration_min+(duration_sec/60)) %>% select(-duration_min, -duration_sec)
dfd <- read_xlsx("PaceCountData_09NOV.xlsx") %>% mutate(duration=duration_min+(duration_sec/60)) %>% select(-duration_min, -duration_sec)
#set dataframe and filter out ALC
df <- dfa %>%
  rbind(dfb) %>%
  rbind(dfc) %>% 
  rbind(dfd) %>%
  mutate(duration=round(duration, 1)) %>% 
  drop_na(pacecount)  %>%
  filter(course!="ALC")  %>%
  mutate(NVG = if_else(course =="RSLC_Cadre" & day_night=="night"|course == "RSLC_Students" & day_night=="night", "Yes", "No")) %>%
  mutate(duration = if_else(course=="RSLC_Cadre" & day_night=="night" & loadcarrying=="load", 0, duration)) %>%
  mutate(wtht_ratio = (weight/height) )

df <- replace_with_na_at(df, .vars="duration", condition = ~.x== 0)
df$loadcarrying <- factor(df$loadcarrying, levels = c("noload", "load"))

count <- df  %>% group_by(ID, course, gender) %>% select(ID, course, gender) %>% unique()
count$gender <- factor(count$gender, levels=c("M", "F"))
count %>% group_by(course, gender) %>% summarise(n=n())


###############################Write total dataset
#write.csv(df, "PaceCountData_21Oct20.csv")

########################################################
#filter out the outliers
#Men outliers
Q_M_day <- df %>% filter(gender=="M", day_night=="day") %>%
     select(pacecount) %>%  quantile(probs=c(.25,.75), na.rm=TRUE)
iqr_M_day <- df %>% filter(gender=="M", day_night=="day") %>%
     select(pacecount) %>% na.omit()
iqr_M_day <- IQR(iqr_M_day$pacecount)
up_M_day <- round(Q_M_day[2]+1.5*iqr_M_day,0)
low_M_day <- round(Q_M_day[1]-1.5*iqr_M_day,0)

Q_M_night <- df %>% filter(gender=="M", day_night=="night") %>% select(pacecount) %>%  quantile(probs=c(.25,.75), na.rm=TRUE)
iqr_M_night <- df %>% filter(gender=="M", day_night=="night") %>% select(pacecount) %>% na.omit()
iqr_M_night <- IQR(iqr_M_night$pacecount)
up_M_night <- round(Q_M_night[2]+1.5*iqr_M_night,0)
low_M_night <- round(Q_M_night[1]-1.5*iqr_M_night,0)

#Female outliers
Q_F_day <- df %>% filter(gender=="F", day_night=="day") %>% select(pacecount) %>%  quantile(probs=c(.25,.75), na.rm=TRUE)
iqr_F_day <- df %>% filter(gender=="F", day_night=="day") %>% select(pacecount) %>% na.omit()
iqr_F_day <- IQR(iqr_F_day$pacecount)
up_F_day <- round(Q_F_day[2]+1.5*iqr_F_day,0)
low_F_day <- round(Q_F_day[1]-1.5*iqr_F_day,0)

Q_F_night <- df %>% filter(gender=="F", day_night=="night") %>% select(pacecount) %>%  quantile(probs=c(.25,.75), na.rm=TRUE)
iqr_F_night <- df %>% filter(gender=="F", day_night=="night") %>% select(pacecount) %>% na.omit()
iqr_F_night <- IQR(iqr_F_night$pacecount)
up_F_night <- round(Q_F_night[2]+1.5*iqr_F_night,0)
low_F_night <- round(Q_F_night[1]-1.5*iqr_F_night,0)

df_outliersremoved_F_day <- df %>% filter(gender=="F", day_night=="day") %>% 
        filter(pacecount >= low_F_day & pacecount <= up_F_day)
df_outliersremoved_F_night <- df %>% filter(gender=="F", day_night=="night") %>% 
        filter(pacecount >= low_F_night & pacecount <= up_F_night)
df_outliersremoved_M_day <- df %>% filter(gender=="M", day_night=="day") %>% 
        filter(pacecount >= low_M_day & pacecount <= up_M_day)
df_outliersremoved_M_night <- df %>% filter(gender=="M", day_night=="night") %>% 
        filter(pacecount >= low_M_night & pacecount <= up_M_night)
df_outliersremoved <- df_outliersremoved_F_day %>% 
     rbind(df_outliersremoved_F_night) %>% 
     rbind(df_outliersremoved_M_day) %>% 
     rbind(df_outliersremoved_M_night)

count2 <- df_outliersremoved  %>% group_by(ID, course, gender) %>% select(ID, gender) %>% unique()
count2 <- count2 %>% group_by( gender) %>% summarise(n=n())
count2_F <- count2 %>% filter(gender=="F")
count2_M <- count2 %>% filter(gender=="M")

#gather data
df2 <- df_outliersremoved %>%  
        gather(pacecount:duration, key=Category, value=measurement)

#weight - height ratio
df_wtht <- df_outliersremoved %>% select(gender, height, wtht_ratio) %>% filter (if_else(gender=="M", height>=66 & height<76,  height>=62 & height<=70 )) %>% unique() %>% summarySE(groupvars = c("height", "gender"), measurevar = "wtht_ratio") %>% mutate(high = (wtht_ratio+sd), low = (wtht_ratio-sd))
df_wtht_M <- df_wtht %>% filter(gender=="M")  
lm_df_wtht_low_M <- lm(low ~ height, df_wtht_M)
lm_df_wtht_high_M <- lm(high ~ height, df_wtht_M)

df_wtht_F <- df_wtht %>% filter(gender=="F")  
lm_df_wtht_low_F <- lm(low ~ height, df_wtht_F)
lm_df_wtht_high_F <- lm(high ~ height, df_wtht_F)

df_wtht %>% select(gender, height, wtht_ratio,  low, high)  %>%
  gather(wtht_ratio:high, key=Category, value=ratio) %>% 
  ggplot (aes(x=height, y=ratio, color=Category)) +
  ylab("weight / height ratio") +
  geom_point() + geom_smooth() + facet_grid(.~gender, scales="free_x")

df_outliersremoved <- df_outliersremoved %>% mutate(wtht_low = if_else(gender=="M", predict(lm_df_wtht_low_M, df_outliersremoved), predict(lm_df_wtht_low_F, df_outliersremoved))) %>% 
  mutate(wtht_high = if_else(gender=="M", predict(lm_df_wtht_high_M, df_outliersremoved), predict(lm_df_wtht_high_F, df_outliersremoved))) %>%
  mutate(wtht_rating = if_else(wtht_ratio <= wtht_low, "low", if_else(wtht_ratio < wtht_high, "normal", "high")))

df_outliersremoved %>% ggplot(aes(x=height, y=pacecount, color=wtht_rating)) + geom_point()  + geom_smooth() + ylim(50, 120) +facet_grid(loadcarrying~gender, scales = "free_x")

df_outliersremoved %>% filter(height >65 & height < 75, day_night =="night") %>% summarySE(measurevar = "pacecount", groupvars = "NVG") 

df2_times <- df2 %>% filter(Category=="duration", measurement>0)
df2_times_summary  <- df2_times  %>%  summarySE(measurevar = "measurement", groupvars = c( "course", "day_night", "loadcarrying" ), na.rm = TRUE )
df2_times_summary_day <- df2_times  %>%  summarySE(measurevar = "measurement", groupvars = c( "day_night" ), na.rm = TRUE ) %>% filter(day_night=="day")
df2_times_summary_night <- df2_times  %>%  summarySE(measurevar = "measurement", groupvars = c( "day_night" ), na.rm = TRUE ) %>% filter(day_night=="night")

df2_times_summary2  <- df2_times    %>% filter(gender=="M") %>%   summarySE(measurevar = "measurement", groupvars = c( "day_night", "loadcarrying" ), na.rm = TRUE )

df_outliersremoved2x <- df_outliersremoved %>%  filter(duration>0) %>%  
     left_join(df2_times_summary2, by =c("loadcarrying", "day_night")) %>%
  mutate(dur_zscore = (duration-measurement)/sd) %>%  
  filter(loadcarrying=="noload", day_night=="night") %>% 
  group_by(ID) %>% 
  summarise(dur_zscore = mean(dur_zscore)) %>% 
  mutate(ExperienceLevel = if_else(dur_zscore <= -.75, "advanced", if_else(dur_zscore <.8, "intermediate", "beginner")))
df_outliersremoved2 <- df_outliersremoved %>% left_join(df_outliersremoved2x) %>% na.omit(ExperienceLevel)

df_outliersremoved2 %>% select(ID, ExperienceLevel, gender) %>% unique() %>% group_by(ExperienceLevel) %>% summarise(n=n()/274)

df_outliersremoved2 %>% summarySE(measurevar = "ruckweight", groupvars = c("ExperienceLevel" ), na.rm = TRUE )

df_outliersremoved2 %>%  filter(loadcarrying=="load") %>% ggplot() + geom_boxplot(aes(x=ExperienceLevel, y=ruckweight)) 
################# Write file with outliers removed
#write.csv (df_outliersremoved, "PaceCountData_21Oct20_outliersremoved.csv")



#Plot of linear regression plots (all categories)
df_outliersremoved %>%  
        ggplot(aes(x=height, y=pacecount, color=day_night))  + 
        geom_point(color="gray") + 
     geom_smooth(method = "lm", aes(linetype=loadcarrying, color=day_night), size=1) + ylab("pacecount") +
     theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
     scale_color_manual(values=c("blue", "tomato")) +
     facet_grid(.~gender, scales = "free_x")
```

## Sample Summary

<div class="blue">
Study sample size: Females n=`r count2_F$n`; Males n=`r count2_M$n`; ~1,500 observations.
</div>
```{r sample summary,  message=FALSE, warning=FALSE}
count %>% group_by(course, gender) %>% summarise(n=n()) %>%
  ggplot(aes(x=reorder(course, n, fun=max), y=n, fill=gender)) + 
  geom_col() + 
  ylab("# of study participants") +
  xlab("")+
  geom_text(aes(label=n), hjust=1) +
  scale_fill_manual(values=c("skyblue", "tomato")) +
  coord_flip()
```

## Key Findings (1 of 2)
 
- ***At Night*** condition has a marked impact on pace count (approx. +4 steps for advanced trainees and +10 steps for beginners).

- ***Height*** displays a moderate negative correlation with pace count (-0.4), suggesting an impact from stride length. 

- ***Load/Ruck weight*** has a marked impact to pace count (~ +4 steps from comparable noload conditions); ruck weights over 35 lbs having increasing impact.

- ***Duration*** appears to be a proxy measure for ***experience***; lower duration, especially at night, suggests greater experience and generally lower pace counts.

## Key Findings (2 of 2)

- ***Gender*** appears to have an impact on pace count, though multi-collinearity with height and "experience" reduces the information gain from this feature. 

- ***Soldier Weight*** strongly correlates with height and has negligible information gain for predicting pace count. 

- ***Night Vision Goggles*** may contribute to a minor reduction (~ -2 steps) in pace counts at night, when compared to trainees without NVGs.  This finding requires further exploration to determine if experience of RSLC Students is confounding.

- ***For further exploration:*** 20% (52/259) record lower pace counts under load, generally at night.  

## Model Recommendation

***Essential Features (R2 = 0.535):***

- Day or Night Condition
- Soldier Height  
- Load/Ruckweight or Noload Condition 

***Added Value (R2 = 0.597):***

- Experience (beginner/intermediate/advanced) 

***Small Information Gain:***

- Gender (R2 = 0.614)  
- Soldier Weight (R2 = 0.615)  


## Correlation plot

```{r correlation, message=FALSE, warning=FALSE}
df_corrplot <- df_outliersremoved2 %>% filter(length == "long") %>% 
        select (pacecount, ruckweight, day_night, height, weight, gender,  ExperienceLevel) %>%
               mutate(day_night = if_else(day_night == "day", 0, 1), 
               gender = if_else(gender == "F", 0,1),
               experience = if_else(ExperienceLevel=="advanced", 2, if_else(ExperienceLevel=="intermediate", 1, 0))) %>% select(-ExperienceLevel)

corrplot(cor(df_corrplot), method="color", order="hclust", type="full", addrect=2, cl.lim=c(-1,1), addCoef.col="black", rect.col="green", diag = FALSE, number.font=1, number.digits = 1, number.cex = .7)
```

## Day/Night & Solider Height

```{r heightimpact,  message=FALSE, warning=FALSE}
df2 %>%  filter(Category=="pacecount", length=="long") %>% 
     ggplot(aes(x=height,y=measurement, color=day_night)) +
     geom_point(alpha=.4) + geom_smooth() +
     scale_color_manual(values = c("green", "blue")) +
     theme(legend.title= element_text(color="black", size=10), legend.position = "right") +
     ylab("pace count") + xlim(58,76) + ylim(55,100)
```

## Load (1 of 2)
- points above dashed line represent pace count ***increase***
```{r load1,  message=FALSE, warning=FALSE}
df2 %>% filter(length=="long", Category=="pacecount") %>% 
        select(course , ID, day_night, loadcarrying, measurement)  %>% 
        spread(key=loadcarrying, value=measurement) %>%  na.omit() %>% 
     ggplot(aes(x=noload, y=load, color=day_night)) + 
     geom_point() + 
     geom_smooth(se=FALSE) + 
     geom_abline(slope = 1, linetype="dashed", color='red') + 
        ylab("pace count UNDER load") + 
        xlab("pace count NO load") +
        scale_color_manual(values = c("lightgreen", "blue")) + 
        ylim(55,85) + 
        xlim(55,85)

# df2 %>% filter(length=="long", Category=="pacecount", day_night=="night") %>% 
#          select(course , ID, day_night, loadcarrying, measurement)  %>% 
#          spread(key=loadcarrying, value=measurement) %>%  na.omit() %>% mutate(delta=(load-noload))  %>% filter(delta<=0) %>% summarySE(measurevar = "delta", groupvars = "course")
```

## Load (2 of 2)
RSLC Student ave ruck = 25 lbs; OCS ave ruck = 50 lbs.
```{r load2, message=FALSE, warning=FALSE}
df_outliersremoved2a <- df_outliersremoved %>% select(-ruckweight, -duration) %>% filter(length!="short") %>% spread(key=loadcarrying, value = pacecount) 

df_outliersremoved2b <- df_outliersremoved %>% select(ID, ruckweight) %>% filter(ruckweight>0) %>% right_join(df_outliersremoved2a) %>% mutate(LoadDelta = (load-noload)) %>% unique()

# df_outliersremoved2b %>% select(course, ID, ruckweight) %>% na.omit() %>% unique() %>%  summarySE(groupvars = "course", measurevar = "ruckweight")

df_outliersremoved2b %>% ggplot(aes(x=height, y=LoadDelta, color=course)) + 
  geom_point() + 
  geom_smooth(se=FALSE) + 
  theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
  ylim(-15,25) +
  scale_color_manual(values = c("gray", "tomato", "red", "green")) +
  geom_hline(yintercept = 0, linetype="dashed", color="black") +
  ylab("individual pace count delta with load") 
```

## Ruck Weight
  
```{r ruckweight1,  message=FALSE, warning=FALSE}
#impact of rucks > 35 lbs on pace count
#df_outliersremoved2b %>% mutate(WeightCategory = ifelse(ruckweight>35, "heavy", "light"))  %>%  summarySE(groupvars = c("WeightCategory") , measurevar = "LoadDelta", na.rm = TRUE)
#correlation
#cor(df_outliersremoved2b  %>% select(ruckweight, LoadDelta) %>% na.omit())

df_outliersremoved2b %>% 
     ggplot(aes(x=ruckweight,y=LoadDelta, color=day_night)) +
     geom_density(aes(x=ruckweight, y=..count..), color="gray",  fill="skyblue", alpha=0.2) +
     geom_point(alpha=.5) + 
     geom_smooth() +  
     geom_hline(yintercept = 0, linetype="dashed", color="black") + 
     scale_color_manual(values = c("green", "blue")) +
     ylab("individual pace count delta with load") + 
     ylim(-5, 24) + 
     xlim(15,55) +
    xlab("ruck weight (lbs)") +
     geom_vline(xintercept = 35, color="red", linetype="dotted", size=1) +
     theme(legend.title= element_text(color="black", size=10), legend.position = "right")+
     labs(caption = "Note: density function drawn for number of observations")
```

## Gender

```{r gender,  message=FALSE, warning=FALSE}
# df2x <- df2 %>% filter(height>62 & height<71) %>% filter(Category=="pacecount", course=="OCS") %>% summarySE(measurevar = "measurement", groupvars = c("gender", "day_night", "height"))
# 
# df2x$height <-factor(df2x$height)

df2  %>% filter(Category=="pacecount") %>%   
  ggplot(aes(x=height, y=measurement, color=gender, group=gender)) + 
  geom_point (alpha=.4) +
  geom_smooth() +
  ylab("pace count") +
  scale_color_manual(values = c("tomato", "skyblue")) +
   theme(legend.title= element_text(color="black", size=10), legend.position = "right")+
  ylim(55,120) 
```

## Solider Weight

```{r weightimpact,  message=FALSE, warning=FALSE}
df2 %>%  filter(Category=="pacecount") %>% mutate(wt_ht=weight/height) %>%  
     ggplot(aes(x=wt_ht,y=measurement, color=gender)) +
     geom_point(alpha=.5) + geom_smooth() +
     geom_hline(yintercept = 0, linetype="dashed", color="red") + 
     scale_color_manual(values = c("tomato", "skyblue")) +
     theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
     ylab("pace count") + xlab("weight to height ratio (lbs/inch)") +
        ylim(55,90)
```

## Night Vision

```{r nightvision1,  message=FALSE, warning=FALSE}
df2 %>% filter(day_night=="night", gender=="M", Category=="pacecount") %>% 
        ggplot(aes(x=height, y=measurement, color=NVG)) +
        geom_point(alpha=.5) +
  ylab("pace count") +
  scale_color_manual(values = c("darkgray", "blue")) +
  theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
  xlim(63,76) +
        geom_smooth(se=FALSE)
```

## Duration (1 of 3)

```{r durationanalysis1,  message=FALSE, warning=FALSE}


df_outliersremoved$day_night <- factor(df_outliersremoved$day_night, levels = c("night", "day"))
df_outliersremoved$loadcarrying <- factor(df_outliersremoved$loadcarrying, levels = c("noload", "load"))

df_outliersremoved %>%  filter (duration>0) %>% ggplot(aes(x=duration, y=pacecount)) +
     geom_point() +
     facet_grid(day_night~.) +
  xlab("time duration (min)") +
  ylab("pace count") +
     geom_smooth() +
     geom_vline(xintercept = df2_times_summary_day$measurement, color="green", linetype = "dashed" ) +
     geom_vline(xintercept = df2_times_summary_night$measurement, color="red", linetype = "dashed" ) +
     labs(caption = "green dashed = mean DAY duration; red dashed = mean NIGHT duration")
```

## Duration (2 of 3)

```{r durationanalysis2,  message=FALSE, warning=FALSE}
df2_times_summary %>%
     ggplot(aes(x=course, y=round(measurement, 2), color=loadcarrying)) +
     geom_point() +
     geom_errorbar(aes(ymin=measurement-ci, ymax=measurement+ci), width=.1 ) +
     coord_flip()  +
     geom_text(aes(label=round(measurement,1)), size=3, vjust=-.4, color="black") +
     ylab("duration on 600m course in minutes") +
     theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
     scale_color_manual(values=c("green", "tomato")) +
     xlab("") + facet_grid(day_night~., scales="free_x") 
```

## Duration (3 of 3)
Does duration explain a level of experience?

```{r durationanalysis3,  message=FALSE, warning=FALSE}


df_outliersremoved2$day_night <- factor(df_outliersremoved2$day_night, levels=c("night", "day"))
df_outliersremoved2$ExperienceLevel <- factor(df_outliersremoved2$ExperienceLevel, levels=c("advanced", "intermediate", "beginner"))

df_outliersremoved2 %>% na.omit(ExperienceLevel) %>% 
     ggplot(aes(x=loadcarrying, y=pacecount, color=ExperienceLevel)) +
     geom_boxplot() +
     stat_summary(aes(label=round(..y..,2)), fun.y="median", geom="text", size=4, color="red", vjust=-.2) +
     facet_grid(day_night~ExperienceLevel) +
     xlab("") +
     scale_color_manual(values = c("green", "skyblue", "tomato")) +
  theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
     ylim(55,95) +
  ylab ("pace count") +
     labs(caption = "advanced = fastest 20% duration; beginner = slowest 20% duration; assessed from noload condition")
```

## Back-up: Pace Count Table
- Working

## Back-up: Experience

```{r experience,  message=FALSE, warning=FALSE}
df_outliersremoved2 %>% ggplot(aes(x=height, y= pacecount, color=ExperienceLevel)) +
  geom_point() + 
  geom_smooth() + 
  facet_grid(day_night~., scales = "free_y") +
     scale_color_manual(values = c("green", "skyblue", "tomato")) +
  theme(legend.title= element_text(color="black", size=10), legend.position = "top") 
```

## Back-up: Precision
```{r experience2,  message=FALSE, warning=FALSE}
df_precision_plot <- df %>% select (course, ID, pacecount) %>% group_by(course, ID) %>% summarise(pacecount_sd = sd(pacecount)) 

df_precision_plot %>% 
  ggplot(aes(x=reorder(course, pacecount_sd,fun=median), y=pacecount_sd)) + 
  geom_boxplot() +
  stat_summary(aes(label=round(..y..,1)), fun.y="median", geom="text", size=3, color="red", vjust=-.2) + ylim(0,15) + xlab("") + ylab ("participant pace count std deviation") +
  labs(caption = "Note: std deviation determined from 4 trials (night/day, load/noload)")
```

## Back-up: What Matters (1 of 2)
Machine Learning with Regression Trees
```{r whatmatters1,  message=FALSE, warning=FALSE}
df_regressiontrees <- df_outliersremoved2 %>%  
     filter(length=="long") 
m <- rpart (pacecount ~ ExperienceLevel + height + day_night + ruckweight   , data = df_regressiontrees)
predict_trees <- predict(m, df_regressiontrees)
regtree_importance <- as.data.frame(m$variable.importance) %>% as.data.frame()
max_importance <- max(regtree_importance$`m$variable.importance`)
regtree_report <- rownames_to_column(regtree_importance, var="Feature") %>%
        mutate(Importance = round(`m$variable.importance`/max_importance,1)) %>% 
        select ( Feature, Importance) %>%  arrange(-Importance)
regtree_report %>% ggplot () + geom_col(aes(x=reorder(Feature, Importance, fun=max), y=Importance), fill= "skyblue") + coord_flip() + xlab("") + ylab("importance in pace count prediction")
```

## Back-up: What Matters (2 of 2)

```{r whatmatters2,  message=FALSE, warning=FALSE}
rpart.plot(m, digits=2, fallen.leaves = TRUE, type=5, extra=101)
```

## Back-up: Duration and day/night Delta

```{r durationanalysis4,  message=FALSE, warning=FALSE}
#Analysis of Impact of Proficiency.
df_comp <- df_outliersremoved %>% filter(loadcarrying=="noload", length=="long") %>% select(ID,course, day_night,pacecount) %>% spread(key=day_night, value=pacecount) %>% na.omit() %>% mutate(day_night_noload_delta=night-day) %>% select(ID, course, day_night_noload_delta)

df_comp_b <- df_outliersremoved %>% filter(loadcarrying=="noload", length=="long") %>% select(ID,course, day_night,duration) %>% spread(key=day_night, value=duration) %>% na.omit() %>% mutate(day_night_noload_delta_duration=night-day) %>% select(ID, course, day_night_noload_delta_duration)

df_comp2 <- df_outliersremoved %>% filter(loadcarrying=="load", length=="long") %>% select(ID, course, day_night,pacecount) %>% spread(key=day_night, value=pacecount) %>% na.omit() %>% mutate(day_night_load_delta=night-day) %>% select(ID, course, day_night_load_delta)

df_comp2_b <- df_outliersremoved %>% filter(loadcarrying=="load", length=="long") %>% select(ID, course, day_night,duration) %>% spread(key=day_night, value=duration) %>% na.omit() %>% mutate(day_night_load_delta_duration=night-day) %>% select(ID, course, day_night_load_delta_duration)

Ave_durationdelta <- df_comp %>% left_join (df_comp2) %>% left_join(df_comp_b) %>% left_join(df_comp2_b) %>% na.omit() %>% mutate(Duration_delta = ((day_night_noload_delta_duration + day_night_load_delta_duration )/2)) %>% select(Duration_delta)
#summary(Ave_durationdelta)

df_comp %>% left_join (df_comp2) %>% left_join(df_comp_b) %>% 
        left_join(df_comp2_b) %>%
        na.omit() %>% mutate(Duration_delta = ((day_night_noload_delta_duration + day_night_load_delta_duration )/2), NightDuration_Delta = if_else(Duration_delta>2.45, "High > 2.5 min", if_else(Duration_delta>1.25, "Average", "Low < 1.25 min"))) %>%
        ggplot(aes(x=day_night_noload_delta, y=day_night_load_delta, color=NightDuration_Delta)) + 
        geom_point(size=2)  +
        ylim(0,25) +
        xlim(0,25) +
        geom_abline(slope = 1, linetype="dashed", color="black", size=1) +
        geom_hline(yintercept = 6, linetype="dotted", color="green", size=1)+
        geom_vline(xintercept = 6, linetype="dotted", color="green", size=1) +
        scale_color_manual("D_N duration delta", values = c("darkgray", "red", "green")) +
        xlab("Day_Night pacecount delta (noload)") +
        ylab("Day_Night pacecount delta (load)") +
        theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
        facet_wrap(course~.)
```

## Back-up: Course Comparison (1 of 2)
Individual results: points expected near dashed line

```{r coursecomparison2,  message=FALSE, warning=FALSE}
df %>% filter(loadcarrying=="noload", day_night=="day") %>% 
        select(course, ID, length, pacecount) %>% 
        spread(key=length, value=pacecount)  %>% 
        na.omit() %>%
        mutate(Delta=abs(long-short)) %>% 
        mutate(Delta_Courses=ifelse(Delta>3, "large(>3)", "small(<=3)")) %>% 
     ggplot(aes(x=short, y=long, color=Delta_Courses)) +
     geom_point(alpha=.7) + 
     geom_abline(slope = 1, linetype="dashed", color='red') + 
     scale_color_manual(values=c("tomato", "skyblue")) +
     theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
     ylab("pace count (600m course)") + 
     xlab("pace count (100m course)") +
     xlim(55,85) + 
     ylim(55,85)
```

## Back-up: Course Comparison (2 of 2)

```{r course comparison1, message=FALSE, warning=FALSE}
#Comparison of short and long course results (day, noload)
df_outliersremoved %>% 
     filter(loadcarrying=="noload", day_night=="day") %>% 
     ggplot(aes(x=height, y=pacecount, color=length)) + geom_point() + 
     geom_smooth() +
     scale_color_manual(values=c("tomato", "green")) +
     theme(legend.title= element_text(color="black", size=10), legend.position = "top") +
         labs(caption = "Note: day and noload conditions")
```